Reviving the Lost Art of Combinator Graph Reduction

### 2022-12-17 experiments/superg/doc/paper/paper.tex
Write on eta optimization.

\usepackage{amsmath}
\DeclareMathOperator{\strict}{strict}
\DeclareMathOperator{\lazy}{lazy}

@@

\textbf{TODO:} Eta optimization.

>>>>

\usepackage{amsmath}
\DeclareMathOperator{\strict}{strict}
\DeclareMathOperator{\lazy}{lazy}
\DeclareMathOperator{\lazyeta}{lazy_\eta}

@@

\subsection{Eta optimization}
Kiselyov covers eta optimization as a change to his OCaml code rather than typing rules.
Our recursive functions are derived from the typing rules, so these changes are not as straightforward to incorporate.
%Previously Kiselyov's representation consisted of elements $C$, $N$ and $W$, representing closed expressions, a used context element, and an ignored context element respectively.
The introduction of the $V$ element means that a context may now be associated with either a closed expression or the top variable (we reuse $V$ as a marker for this).
Our implementation uses a tagged union to define a type that can be either an expression or the top variable.
For the compilation function, we only need to change the definition of $\lazyeta \ z$ and add an extra case for $\lazyeta \ \lambda e$:

\begin{equation*}
    \begin{array}{l l l l}
        \lazyeta & z         & = & \emptyset \models V                                                     \\
        \lazyeta & s \ e     & = & \Gamma,f \models c                                                      \\
                 &           &   & \text{where } (\Gamma \models c) := \lazyeta e                          \\
        \\
        \lazyeta & \lambda e & = & \emptyset \models I                                                     \\
                 &           &   & \text{where } (\emptyset \models V) := \lazyeta e                       \\
        \lazyeta & \lambda e & = & \emptyset \models K \; c                                                \\
                 &           &   & \text{where } (\emptyset \models c) := \lazyeta e                       \\
        \lazyeta & \lambda e & = & (\emptyset \models K) \coprod (\Gamma \models c)                        \\
                 &           &   & \text{where } (\Gamma,f \models c) := \lazyeta e                        \\
        \lazyeta & \lambda e & = & \Gamma \models K \; c                                                   \\
                 &           &   & \text{where } (\Gamma,t \models c) := \lazyeta e                        \\
        \\
        \lazyeta & e_1 \ e_2 & = & (\Gamma \models ((\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)) \\
                 &           &   & \text{where } (\Gamma_1 \models c_1) := \lazyeta e_1                    \\
                 &           &   & \text{where } (\Gamma_2 \models c_2) := \lazyeta e_2                    \\
                 &           &   & \text{where } \Gamma := \Gamma_1 \sqcup \Gamma_2                        \\
    \end{array}
\end{equation*}

\textbf{TODO: } Check this, possibly need to take (C d, V) = N (C d)
The context merge function $\sqcup$ is unchanged, but we do have some new cases for the semantic function:

\begin{equation*}
    \arraycolsep=1.4pt
    \begin{array}{l l l l l}
        % Added for eta optimization
        (\Gamma_1,f \models c_1) & \coprod & (\emptyset \models V)    & = & c_1                                                      \\
        (\emptyset \models V)    & \coprod & (\Gamma_2,f \models c_2) & = & (\emptyset \models CI) \coprod (\Gamma_2 \models c_2)    \\

        (\Gamma_1,t \models c_1) & \coprod & (\emptyset \models V)    & = & (\Gamma_1 \models (s \coprod l)) \coprod i               \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } s := (\emptyset \models S )                \\
                                 &         &                          &   & \text{where } i := (\emptyset \models I )                \\
        (\emptyset \models V)    & \coprod & (\Gamma_2,t \models c_2) & = & (\emptyset \models SI) \coprod (\Gamma_2 \models c_2)    \\
        (\emptyset \models c_1)  & \coprod & (\emptyset \models V)    & = & c_1                                                      \\
        (\emptyset \models V)    & \coprod & (\emptyset \models c_2)  & = & CI c_2                                                   \\
        \\

        % Original
        (\emptyset \models c_1)  & \coprod & (\emptyset \models c_2)  & = & c_1 \ c_2                                                \\
        \\
        (\emptyset \models c_1)  & \coprod & (\Gamma_2,t \models c_2) & = & (\emptyset \models B c_1) \coprod (\Gamma_2 \models c_2) \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\emptyset \models C C c_2) \coprod (\Gamma \models c_1) \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (s \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } s := (\emptyset \models S )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        % Lazy
        (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models c_1) \coprod (\Gamma_2 \models c_2)    \\
        \\
        (\Gamma_1,f \models c_1) & \coprod & (\Gamma_2,t \models c_2) & = & (\Gamma_1 \models (b \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } b := (\emptyset \models B )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        (\Gamma_1,t \models c_1) & \coprod & (\Gamma_2,f \models c_2) & = & (\Gamma_1 \models (c \coprod l)) \coprod r)              \\
                                 &         &                          &   & \text{where } c := (\emptyset \models C )                \\
                                 &         &                          &   & \text{where } l := (\Gamma_1 \models c_1 )               \\
                                 &         &                          &   & \text{where } r := (\Gamma_2 \models c_2 )               \\
        \\
        (\emptyset \models c_1)  & \coprod & (\Gamma_2,f \models c_2) & = & (\emptyset \models c_1) \coprod (\Gamma_2 \models c_2)   \\
        \\
        (\Gamma_1,f \models c_1) & \coprod & (\emptyset \models c_2)  & = & (\Gamma_1 \models c_1) \coprod (\emptyset \models c_2)   \\
    \end{array}
\end{equation*}
